---
title: "p8105_hw3_hx2264"
author: "Hongyao Xie"
date: "October 8, 2018"
output: html_document
---

## Problem 1


```{r echo = FALSE}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
library(tidyverse)
library(ggridges)
library(patchwork)
library(dplyr)
library(forcats)
```


```{r}
# Load the data and use appropriate variable names
data(brfss_smart2010)
brfss <- janitor::clean_names(brfss_smart2010)

# Focus on the "Overall Health" topic
overall_health_index <- which(brfss[,"topic"] == "Overall Health")
brfss <- brfss[overall_health_index,]

# Include only responses from "Excellent" to "Poor"
brfss <- filter(brfss, response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor")

# Organize responses as a factor taking levels from “Excellent” to “Poor”
brfss[, "response"] <- as.factor(unlist(brfss[, "response"]))
```

**In 2002, which states were observed at 7 locations?**

```{r}
index_2002 <- which(brfss$year == 2002)
brfss_2002 <- brfss[index_2002,]
unique_location <- unique(brfss_2002[,"locationdesc"])
location_by_state <- as.factor(substr(unique_location$locationdesc, 1, 2))
names(which(summary(location_by_state) == 7))
```

3 states were observed at 7 locations: `r names(which(summary(location_by_state) == 7))`.

**Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.**

```{r}
brfss %>%
  group_by(year, locationabbr) %>%
  summarize(uniq = length(unique(locationdesc))) %>%
  ggplot(aes(x = year, y = uniq, color = locationabbr)) + 
    geom_line() +
    labs(
      title = "Spaghetti plot of observations in each state from 2002 to 2010",
      x = "Year",
      y = "Number of observations"
    ) +
  theme(legend.position = "right")
```

**Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.**

```{r}
brfss %>%
  filter(., locationabbr == "NY") %>%
  filter(., year == 2002 | year == 2006 | year == 2010) %>%
  filter(., response == "Excellent") %>%
  group_by(year) %>%
  summarize(mean = mean(data_value),
            sd = sd(data_value)) %>%
  knitr::kable(digits = 1)
```

**For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.**

```{r}
#make a five-panel plot
brfss %>% 
  mutate(., response_ordered = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>% 
  group_by( year, locationabbr, response_ordered) %>% 
  summarise(mean = mean(data_value)) %>% 
  ggplot() +
  geom_point(aes(x = year, y = mean),na.rm = TRUE) + 
  labs(
    title = "Five-panel plot of the distribution of state-level averages over time",
    x = "Year",
    y = "Averages"
  ) +
  facet_grid(~response_ordered)
```


## Problem 2

```{r}
library(p8105.datasets)
data("instacart")

itc <- instacart %>%
  janitor::clean_names()
``` 

**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
nrow(unique(itc[,"aisle"]))
names(which.max(table(itc[,"aisle"])))
```

There are `r nrow(unique(itc[,"aisle"]))` aisles, out of which `names(which.max(table(itc[,"aisle"])))` is the most items ordered from.

**Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.**

```{r}

```
















